---
title: "**Introduction to Statistics and Computation with Data: Worksheet 4, Part 2**"
author: "Daibik Barik (Bmat2316)"
date: "2024-02-16"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE}
set.seed(2024)
```

\section{1.}
```{r}
x = c(1,2,3,4,5,6)
probx= c(1/4,1/8,1/8,1/8,1/8,1/4)
F16=sample(x, size=1500, replace=T, prob=probx)
```

\subsection{(a)}
The lines x = c(1,2,3,4,5,6) and probx= c(1/4,1/8,1/8,1/8,1/8,1/4) create vectors for a dice roll and $6$ probabilities respectively.\
The line F16=sample(x, size=1500, replace=T, prob=probx) associates the probability vector 'probx' with the 'x' vector and generates $1500$ samples with replacement (replace=T) according to the given probability of each roll.

\subsection{(b)}
```{r}
mean(F16)
var(F16)
```

We know, Range of sample $\in$ (mean $-$ $\sqrt{\text{var}}$, mean $+$ $\sqrt{\text{var}}$)

```{r}
alpha = mean(F16)-sqrt(var(F16))
beta = mean(F16)+sqrt(var(F16))
alpha; beta
```

$$ \therefore \text{Range(F16)}= (\text{alpha},\; \text{beta}) $$

\subsection{(c)}
We calculate the true mean and variance separately.
$$\text{Mean}(E[X])= \sum_{i=1}^6 X_i \cdot P[X_i] \quad\text{and}\quad \text{Variance}= E[X^2] -(E[X])^2 $$


```{r}
mean_x=0
for(i in (1:6)){
  mean_x= mean_x+ x[i]*probx[i]
}
mean_x
```
```{r}
mean_x2=0
for(i in (1:6)){
  mean_x2= mean_x2+ (x[i])^2*probx[i]
}
mean_x2 - (mean_x)^2
```
$$ \therefore \text{Mean}= 3.5 \quad\text{and}\quad \text{Variance} = 3.75$$
which are quite close to the mean and variance generated by the sample.
\newpage


\section{2.}
```{r}
b1 = rbinom(100,10,0.5)
b2 = rbinom(100,10,0.25)
b3 = rbinom(100,10,0.75)
```

\subsection{(a)}
rbinom(n, size, prob) \ 
It generates $n$ number of random values of given probability(prob) from a binomially distributed sample of given size.

\subsection{(b)}
```{r}
mean(b1); var(b1)
mean(b2); var(b2)
mean(b3); var(b3)
```

For binomial distribution, 
$$ \text{Mean}(E[X])= Np \quad\text{and}\quad \text{Variance}= Np(1-p)$$

```{r}
mean_b1=  10*0.5
mean_b2=  10*0.25
mean_b3=  10*0.75
mean_b1; mean_b2; mean_b3
```
```{r}
var_b1= 10*0.5*(1-0.5)
var_b2= 10*0.25*(1-0.25)
var_b3= 10*0.75*(1-0.75)
var_b1; var_b2; var_b3
```
which are again, quite close to the mean and variance calculated from the sample.

\subsection{(c)}
```{r fig.height=3, fig.width=3}
library(ggplot2)
df1=data.frame(b1)
p11= ggplot(df1) + geom_histogram(mapping=aes(x=b1), color="black", fill="NA", binwidth=1)
p21= ggplot(df1) + geom_histogram(mapping=aes(x=b1, y=after_stat(density)), color="black", fill="NA", binwidth=1)
p11; p21
```

\subsubsection{i.}
The first piece of code('p11') generates a histogram plot utilizing data from the 'df1' data frame. It extracts random samples from 'b1' to create the data frame. The x-axis represents the values of the 'b1' variable, with histogram bars having black outlines and transparent fills. Each bin has a width of $1$ unit.

Conversely, the second code('p21') segment also produces a histogram plot using data from df1. On the x-axis, the values of the 'b1' variable are depicted, and the density of the histogram bars is computed instead of the count. Similar to the first code, the histogram bars have black outlines and transparent fills, and each bin retains a width of $1$ unit.

\subsubsection{ii.}
For b2,
```{r fig.height=3, fig.width=3}
df2=data.frame(b2)
p12= ggplot(df2) + geom_histogram(mapping=aes(x=b2), color="black", fill="NA", binwidth=1)
p22= ggplot(df2) + geom_histogram(mapping=aes(x=b2, y=after_stat(density)), color="black", fill="NA", binwidth=1)
p12; p22
```

For b3,
```{r fig.height=3, fig.width=3}
df3=data.frame(b3)
p13= ggplot(df3) + geom_histogram(mapping=aes(x=b3), color="black", fill="NA", binwidth=1)
p23= ggplot(df3) + geom_histogram(mapping=aes(x=b3, y=after_stat(density)), color="black", fill="NA", binwidth=1)
p13; p23
```

\subsubsection{iii.}
These $3$ plots give us the density and the PDF of all the distributions for easy comparison.

\subsection{(d)}
```{r fig.height=4, fig.width=4}
library(ggplot2)
density = function(x,a,s){ (1/((2*pi)^(0.5)*s ))* exp(-(x-a)^2/(2*s^2))}
df1=data.frame(b1)
p21= ggplot(df1) + geom_histogram(mapping=aes(x=b1, y=..density..), color="black", fill="NA", binwidth=1) + xlim(0,10) + 
  geom_function(fun=density, args=list(a=5,s=(2.5)^(0.5)))
p21
```

\subsubsection{i.}
$\int_3^6 \text{density}(x,5,\sqrt{2.5}) dx$ approximates to $0.67$.\
This denotes that the our distribution sample lies between 3 and 6 with the probability of $0.67$.

\subsubsection{ii.}
The area under the graph from $3$ to $7$ is $0.898$
$$ \Phi(b)-\Phi(a)= 0.898 $$
I guess, $a = -1.273$ and $b= 3.828$

\subsubsection{iii.}
For b2,
```{r fig.height=4, fig.width=4}
density = function(x,a,s){ (1/((2*pi)^(0.5)*s ))* exp(-(x-a)^2/(2*s^2))}
df2=data.frame(b2)
p22= ggplot(df2) + geom_histogram(mapping=aes(x=b2, y=..density..), color="black", fill="NA", binwidth=1) + xlim(0,10) + 
  geom_function(fun=density, args=list(a=5,s=(2.5)^(0.5)))
p22
```
\newpage

For b3,
```{r fig.height=4, fig.width=4}
density = function(x,a,s){ (1/((2*pi)^(0.5)*s ))* exp(-(x-a)^2/(2*s^2))}
df3=data.frame(b3)
p23= ggplot(df3) + geom_histogram(mapping=aes(x=b3, y=..density..), color="black", fill="NA", binwidth=1) + xlim(0,10) + 
  geom_function(fun=density, args=list(a=5,s=(2.5)^(0.5)))
p23
```

No we won't get the same result as the dataframes in use are based on different distributions and are hence different.
