---
title: "Assignment 7.1 LM&R"
author: "Daibik Barik (Bmat2316)"
date: "2025-03-30"
output: pdf_document
---

**Table 1: Production Rates with Replicates**

| Reagent | Catalyst I | Catalyst II | Catalyst III |
|---------|------------|-------------|---------------|
| A       | 4, 6       | 11, 7       | 5, 9          |
| B       | 6, 4       | 13, 15      | 9, 7          |
| C       | 13, 15     | 15, 9       | 13, 13        |
| D       | 12, 12     | 12, 14      | 7, 9          |


```{r}
```


\section*{Q1.}

```{r}
#Generate the Linear Regression Model
model <- lm(Y ~ X1 + X2 + X3, data = data)
summary(model)
```
As we only need the normal equation, we use the coefficients. If the amount of energy is $Y$, and the varuious variables are given by $X_1$ (Area), $X_2$ (Age), $X_3$ (No. of guest rooms); then the equation is- 

$$ Y = a_0 + a_1(X_1- \overline{X_1}) + a_2(X_2- \overline{X_2}) + a_3(X_3- \overline{X_3}) $$
Getting the values of $\overline{X_1}, \overline{X_2}, \overline{X_3}$ from the dataset, we modify the equation to- 

$$ Y = b_0 + \sum_{i=1}^3 b_iX_i \text{ whose coefficients are given by}$$

```{r}
#Fetching the values from the model
coefficients(model)
b0 = unname(coefficients(model)["(Intercept)"])
b1 = unname(coefficients(model)["X1"])
b2 = unname(coefficients(model)["X2"])
b3 = unname(coefficients(model)["X3"])
```


\section*{Q2.} 

For _Goodness of Fit_ measure, $R^2$ and $p$ values are required.
$R^2$ is calculated as  $$ 1-R^2 = \frac{SS_{pred}}{SS_{obs}} = \frac{\sum_{i=1}^{19} (Y-\hat Y)^2}{\sum_{i=1}^{19} (Y-\overline Y)^2} $$

```{r}
SS_obs = sum((data$Y - mean(data$Y))^2) 
#for observed deviation, we just subtract the true mean the Y's and calculate
predicted_Y = b0 + b1 * data$X1+ b2 * data$X2+ b3 * data$X3 
SS_pred = sum((data$Y- (predicted_Y) )^2) 
#for predicted deviation, we calculate the predicted Y value 
# from our regression eqn and use it for calculation
R2= 1 - SS_pred/ SS_obs 
print(R2)
```
which matches the evaluated value in R,

```{r}
summary(model)$r.squared
#The value calculated in R
```
which says that $80.15455\%$ of data in the dataset fits into our predicted regression model.

**~ ALITER ~**

The multiple correlation coefficient of $Y$ (or $X_4$) is given by $Corr(\hat{X4},X4)$, and hence gives us the relation between the observed and the predicted data. 

$$ m_4 = \sqrt{1 - \frac{|R|}{R_{44}}} $$

```{r}
data = data.frame(data)
colnames(data)[colnames(data)=='Y'] = 'X4'
data = data[,c('Hotel','X1','X2','X3','X4')]
subdata = subset(data,select=c(X1=X1,X2=X2,X3=X3,X4=X4))
R = cor(subdata)
m4 = sqrt(1-det(R)/det(R[-4,-4]))
print(m4)
```

The multiple correlation coefficient is $89.5\%$, which is very high. This indicates that the model serves as a good fit to the data.



\section*{Q3.}

Using the values given, we substitute them in R-

```{r}
Y = b0 + b1 * 37+ b2 * 11+ b3 * 215.3
#We are using the constants that we defined previously
print(Y)
```
and get the energy consumption Y as 6234010 kWh.


\section*{Q4.}

The partial correlation between $X_2$ (age) and $Y$ (energy consumption) gives us an idea about their impact on each other. We do it in R-

```{r}
library(ppcor)
colnames(data)[colnames(data)=='X4'] = 'Y'
partial_corr <- pcor(data[, c("Y", "X2", "X1", "X3")]) 
#Calculate the partial correlation matrix
partial_corr_value <- partial_corr$estimate["Y", "X2"] 
#Extract the partial correlation between Y and X2
print(partial_corr_value) 
```
Since the value is a small positive number, increase in age of Hotel implies a slight increase in its energy consumption.

**~ ALITER ~**

We use the formula 
$$ r_{24} = \frac{-R_{24}}{\sqrt{R_{22} \times R_{44}}} $$
in R.

```{r}
r_24 = -1*det(R[-2,-4])/sqrt(det(R[-2,-2])*det(R[-4,-4]))
r_24
```
which matches our previous result.


\section*{Q5.}

We can easily form a regression model by making the relationships **non-linear**, possibly a polinomial regression with squares and cubic terms.

Say the model (non-linear) be-
$$ Y \sim X_1 + X_2^2 + X_3^3 $$

```{r}
model_poly <- lm(Y ~ X1 + I(X2^2)+ I(X3^3), data = data)
summary(model_poly)
```
Since the $R^2 = 0.7853$, which resembles about $78.5\%$ data fit into our model, so it's not a bad one! 
