---
title: "**Introduction to Statistics and Computation with Data: HomeWork 8**"
author: "Daibik Barik (Bmat2316)"
header-includes:
    - \usepackage{bm, amsmath, amssymb}
date: "2024-03-20"
output: pdf_document
---

\section*{1. }
Constructing a confidence interval for the given data, 

Let $X_1, X_2, \dots, X_{n=8} = 300,500,300,700,300,500,300,700$ respectively.
\[\overline X = \frac{300+500+300+700+300+500+300+700}{8} = \frac{3600}{8} = 450\]
It is given that each $X_i \sim \mathcal{N}(\mu, \sigma^2)$, where $\sigma^2 = 100$
\[\implies K =\frac{\sqrt{n}(\bar{X} - \mu)}{\sigma} \sim \mathcal{N}(0,1)\]
To construct a $95\%$ confidence interval, 
we need to find '$z$' such that 
\[P\left[\mu \in \left(\bar X - \frac{z\sigma}{\sqrt{n}},\bar X + \frac{z\sigma}{\sqrt{n}}\right)\right] = 0.95\]

\[P\left[\mu \in \left(450 - \frac{z\sqrt{100}}{\sqrt{8}},450 + \frac{z\sqrt{100}}{\sqrt{8}} \right) \right] = 0.95\]

\[P\left[\frac{\sqrt{2}(\mu - 450)}{5} \in \left( - z,+ z \right) \right] = 0.95\]
\[\implies P[-z< K < z] = 0.95, \text{ where } K \sim \mathcal{N}(0,1) \]
\[\implies \Phi(z) - \Phi(-z) = 0.95\]
\[\implies \Phi(z) - (1 - \Phi(z)) = 0.95\]
\[\implies \Phi(z) = 0.975\]
```{r}
qnorm(0.975)
```
\[\implies z = 1.959964 \approx 1.96\]
Hence the interval we get is
\[\left(450 - \frac{(1.96)\sqrt{100}}{\sqrt{8}},450 + \frac{(1.96)\sqrt{100}}{\sqrt{8}} \right)\]
\[ = \left( 450 - 6.93,450 + 6.93 \right)\]
\[ = \left( 443.07,456.93\right)\]

\section*{2. }

\subsection*{(a). }
$X_1, ..., X_n$ are iid $Uniform(0,\theta)$
\[L(\theta| X_1, ..., X_n) = \prod_{i=1}^n\frac{1}{\theta} = \theta^{-n}\]
\[\log(L(\theta| X_1, ..., X_n)) = -n\cdot \log(\theta)\]

Differentiating on both sides, 
\[\frac{d\log(L(\theta| X_1, ..., X_n))}{d\theta} = \frac{1}{L(\theta| X_1, ..., X_n)}\frac{-n}{\theta}\]
\[\implies \frac{-n}{\theta} \rightarrow 0\]

This happens when $\theta$ takes the maximum possible value,which is $Max\{X_1, ..., X_n\}$

\subsection*{(b). }
$\hat{\theta} = Max\{X_1, ..., X_n\}$ and we have to find the distribution of $T = \frac{ \hat{ \theta }}{ \theta }$

\[P\left[ \frac{\hat{\theta}}{\theta} \leq t\right] = P[\hat{\theta} \leq t\theta]\]
\[ = P[Max\{X_1, ..., X_n\} \leq t\theta]\]
\[ = P[X_1 < t\theta] \cap ... \cap P[X_n < t\theta]\]
\[ = \left(\frac{t\theta}{\theta}\right)^n\]
\[ = t^n\]

\subsection*{(c). }


\section*{3. }
\[\frac{\sqrt{n}(\overline{X} - \mu)}{S} \sim t_{n-1}\]
\[K = \frac{\sqrt{19}(12 - \mu)}{6.3} \sim t_{18}\]

To construct a 95% confidence interval for the true mean, 
\[P\left[\mu \in \left(12 -\frac{z\cdot6.3}{\sqrt{19}},12 +\frac{z\cdot6.3}{\sqrt{19}} \right)\right] = 0.95\]
\[P[K \in (-z,z) = 0.95]\]
```{r}
qt(0.975, 18)
```
$\implies z = 1.734064 \approx 1.7$ 

So the 95% confidence interval is, 
\[\left(12 -\frac{2.1\times 6.3}{\sqrt{19}},12 +\frac{2.1\times6.3}{\sqrt{19}}\right)\]
\[= \left(8.96483,15.03517\right)\]

\section*{4. }

Given, $X_i \sim Bernoulli(p), \text{ } \forall \text{ } x \in \{1,2, .. n\}$
We estimate p as $\hat{p} = \frac{1}{n}\sum_1^n X_i$
For the bernoulli distribution, the standard deviation is estimated by, 
\[\sqrt{\hat{p}(1-\hat{p})}\] 
Now,the Central Limit Theorem says, 
\[\frac{\sqrt{n}(X - \hat{p})}{\sqrt{\hat{p}(1-\hat{p})}} \sim \mathcal{N}(0,1)\]
So, to find the 95% confidence interval we have to find z such that, 
\[P\left[K = \frac{\sqrt{n}(X - \hat{p})}{\sqrt{\hat{p}(1-\hat{p})}} \in (-z,z) \right] = 0.95\]
\[\implies P[-z < K < z] = 0.95, \text{ where } K \sim N(0,1) \]
\[\implies \Phi(z) - \Phi(-z) = 0.95\]
\[\implies \Phi(z) = 0.975\]
```{r}
qnorm(0.975)
```
\[\implies z = 1.959964 \approx 1.96\]
So, we say that the 95% confidence interval for p is 
\[\left(\hat{p} - 1.96\sqrt{ \frac{\hat{p}(1-\hat{p})}{n}}, \hat{p} + 1.96\sqrt{ \frac{\hat{p}(1-\hat{p})}{n}}\right)\]

\section*{5. }
\subsection*{(a). }
Given, $X_i \sim \mathcal{N}(0,\sigma^2), \text{ } \forall \text{ } i \in \{1,2, .. n\}$
\[\implies \frac{X_i}{\sigma} \sim \mathcal{N}(0,1)\]
\[\implies \left(\frac{X_i}{\sigma}\right)^2 \sim \Gamma\left(\frac{1}{2}, \frac{1}{2}\right)\]
Now, by the additivity of the $\Gamma$ distribution, 
\[\implies \left(\sum_{i=1}^n\left(\frac{X_i}{\sigma}\right)^2\right) \sim \Gamma\left(\frac{n}{2}, \frac{1}{2}\right) \sim \chi_n^2\]
Hence, $\left(\sum_{i=1}^n\left(\frac{X_i}{\sigma}\right)\right)^2$ follows the chi-squared distribution with n degrees of freedom.

\newpage
\subsection*{(b). }
```{r}
alpha = 0.95
n = 100
lower_value = qchisq((1-alpha)/2,n)
upper_value = qchisq((1+alpha)/2,n)
cat("The 95% confidence interval for the chi-squared distribution 
            with 100 degrees of freedom is (", lower_value, ",", upper_value , ")")
```
\subsection{(c). }
```{r}
alpha = 0.95
n = 100
lower_value = qchisq(1-alpha,n)
upper_value = qchisq(alpha,n)
cat("The 95% confidence interval for the chi-squared distribution 
            with 100 degrees of freedom is either", "(", 0, ",", upper_value , ")", 
            ("or"), "(",lower_value, ",", Inf , ")" )
```




