---
title: "**Introduction to Statistics and Computation with Data: HomeWork 3**"
author: "Daibik Barik (Bmat2316)"
date: "2024-01-31"
output: pdf_document
---

\section{1}

``` r
install.packages("forcats")                                                
install.packages("gapminder")
```

```{r setup, echo=FALSE}
install.packages(c("gapminder", "forcats"),
  repos = "http://cloud.r-project.org"
)
```

\newpage

\section{2}

```{r}
library(tidyverse)
library(gapminder)
str(gapminder$continent)
```

Concisely exhibit the internal composition of an R object using a diagnostic function, serving as an alternative to the *summary()*. So here it displays about the 5 factors of *gapminder*, which are also continents named "Africa", "America" etc.

```{r}
levels(gapminder$continent)
```

The *levels()* grants access to the levels attribute of a variable. Here the levels are named after different continents like "Africa", "Americas", etc.

```{r}
nlevels(gapminder$continent)
```

Return the number of levels which its argument has. Here it has exactly 5.

```{r}
class(gapminder$continent)
```

Several R objects contain a class attribute, which is a character vector that provides the names of the classes from which the object inherits. \newpage

\section{3}

```{r}
gapminder %>% count(continent)
```

*count()* lets you quickly count the unique values of frequencies of one or more variables.

```{r}
fct_count(gapminder$continent)
```

*fct_count()* Count entries in a factor, here it gives same output because number of entries in factors and frequencies of the variables are same. \newpage

\section{4}

```{r}
nlevels(gapminder$country)
```

Prints the number of levels in the country column of the gapminder dataset.

```{r}
h_countries <- c("Egypt", "Haiti", "Romania", "Thailand", "Venezuela")
h_gap <- gapminder %>% filter(country %in% h_countries)
nlevels(h_gap$country)
```

It prints the number of levels in the country column, and **since the other levels are still present with 0 entries**, it shows the total number of levels same as before.

```{r}
h_gap_dropped <- h_gap %>%droplevels()
nlevels(h_gap_dropped$country)
```

*droplevels()* removes the unused levels from a factor. So, now if we count number of levels, it returns 5.

```{r}
h_gap$country %>% fct_drop() %>%levels()
```

*fct_drop()* drops the unused levels in the factor, and *levels()* print the remaining levels, by default in alphabetical order. \newpage

\section{5}

```{r}
gapminder$continent %>% levels()
```

Prints the levels in the continent column of Gapminder dataset.

```{r}
gapminder$continent %>% fct_infreq() %>% levels()
```

Prints the levels in the decreasing order of the number of observations.

```{r}
gapminder$continent %>%fct_infreq() %>%fct_rev() %>%levels()
```

Prints the levels in the increasing order of the number of observations (**opposite to the order made by *fct_infreq***)

```{r}
ggplot(data= gapminder) + 
  geom_bar(mapping= aes(x= continent)) +
  coord_flip()
```

```{r}
ggplot(data= gapminder) + 
  geom_bar(mapping= aes(x= fct_rev(fct_infreq(continent)))) +
  coord_flip() +xlab("continent")
```

\newpage

\section{6}

```{r}
fct_reorder(gapminder$country, gapminder$lifeExp) %>%levels() %>%head()
```

This sorts the levels(countries), acc to the ascending order of median life expectancy, the lowest being Sierra Leone at 37.55

```{r}
fct_reorder(gapminder$country,gapminder$lifeExp,min)%>%levels()%>%head()
```

This sorts the levels(countries), acc to the ascending order of min life expectancy ever recorded, the lowest being Rwanda at 23.6

```{r}
fct_reorder(gapminder$country, gapminder$lifeExp,.desc=TRUE)%>%levels() %>% head()
```

This sorts the levels(countries), acc to the descending order of median life expectancy, the higest being Iceland at 76.55 \newpage

```{r}
gap<-subset(gapminder,year==2007 & continent=="Asia")
ggplot(gap)+geom_point(aes(lifeExp,country))
```

\newpage

```{r}
gap<-subset(gapminder,year==2007 & continent=="Asia")
ggplot(gap)+geom_point(aes(lifeExp,forcats::fct_reorder(country, lifeExp)))+ ylab("continent")
```

\newpage

Without using *fct_reorder()*,

```{r}
ggplot(gap)+geom_point(aes(lifeExp,forcats::fct_infreq(country,lifeExp)%>%fct_rev()))+
ylab("continent")
```

\newpage

\section{7}

```{r}
gapsub<-subset(gapminder,country=="Egypt"|country=="Haiti"|country=="Romania"|
country=="Thailand"|country=="Venezuela")
ggplot(gapsub)+geom_line(aes(year,lifeExp,color=country))
```

\newpage

```{r}
ggplot(gapsub)+
geom_line(aes(year,lifeExp,
color=forcats::fct_relevel(country,c("Venezuela","Romania","Egypt",
                                     "Thailand","Haiti"))))+ 
  labs(color="country")
```

\newpage

```{=tex}
\section{8}
\subsection{10.8}
```
```{r}
gapcode <- subset(gapminder,country=="Australia"|country=="Korea, Dem. Rep."|
country=="Korea, Rep.")
gapcode$country %>% droplevels()
```

```{r}
gapcode$country=gapcode$country %>% fct_recode("Oz" = "Australia",
"North Korea" = "Korea, Dem. Rep.",
"South Korea"="Korea, Rep.") %>% droplevels()
```

\newpage

\subsection{10.9}

```{r}
df1 <- subset(gapminder,country=="United States"|country=="Mexico"&year > 2000) %>% droplevels()
df2 <- subset(gapminder,country=="France"|country=="Germany"&year > 2000) %>% droplevels()
```

```{r}
levels(df1$country)
levels(df2$country)
```

```{r}
fct_c(df1$country,df2$country)
```

```{r}
bind_rows(df1,df2)
```

\newpage

```{r}
rbind(df1,df2)
```
