---
title: "**Introduction to Statistics and Computation with Data: HomeWork 4**"
author: "Daibik Barik (Bmat2316)"
date: "2024-02-10"
output: pdf_document
---

\section{1.}

Let $X_1, X_2, \dots ,Xn$ be i.i.d. samples from Normal ($\mu$, $\sigma^2$). Then Likelihood is defined as 
$$ L(X_1, X_2, \dots, X_n| \mu, \sigma^2) = \prod_{i=1}^n \frac{e^\frac{-(X_i-\mu)^2}{2\sigma^2}}{\sigma\sqrt{2\pi}}$$

$$= \frac{e^{\sum_{i=1}^{n}\frac{-(X_i-\mu)}{2\sigma^2}}}{(\sigma\sqrt{2\pi})^n} $$

For simplicity, we consider logarithm. Taking log with base $e$ on both sides and simplifying, we get

$$ l(X_1,X_2,\dots,X_n|\mu,\sigma^2) := ln(L(X_1, X_2, \dots, X_n| \mu, \sigma^2)) $$ 
$$ = -\frac{n}{2}ln(2\pi)-\frac{n}{2}ln(\sigma^2)-\frac{1}{2\sigma^2}\sum_{i=1}^{n}(X_i-\mu)^2 $$

Maximum likelihood estimate for $\mu$ is found by differentiating the expression w.r.t. to the mean ($\mu$).

Let 
$$ \frac{\partial}{\partial\mu}l(X_1,X_2,\dots,X_n|\mu,\sigma^2)= 0 $$ $$ \implies \frac{\partial}{\partial\mu}\left(-\frac{n}{2}ln(2\pi)-\frac{n}{2}ln(\sigma^2)-\frac{1}{2\sigma^2}\sum_{i=1}^{n}(X_i-\mu)^2\right) = 0$$ $$ \implies \frac{1}{\sigma^2}\sum_{i=1}^{n}(X_i-\mu) = 0 $$ $$ \frac{1}{\sigma^2} \left(\sum_{i=1}^{n}X_i-n\mu \right) $$
\newpage

which is equal to zero only when, for some $\hat{\mu}$,

$$ \sum_{i=1}^{n}X_i-n\hat{\mu} = 0 $$ 
$$ \boxed{\hat{\mu} = \frac{1}{n}\sum_{i=1}^{n}X_i} $$

Similarly, for the variance($\sigma^2$), we consider

Let
$$ \frac{\partial}{\partial\sigma^2}l(X_1,X_2,\dots,X_n|\mu,\sigma^2)= 0 $$ 
$$ \implies \frac{\partial}{\partial\sigma^2}\left(-\frac{n}{2}ln(2\pi)-\frac{n}{2}ln(\sigma^2)-\frac{1}{2\sigma^2}\sum_{i=1}^{n}(X_i-\mu)^2\right) = 0 $$ 
$$ \implies -\frac{n}{2\sigma^2}-\left[\frac{1}{2}\sum_{i=1}^{n}(X_i-\mu)^2 \right]\frac{\partial}{\partial\sigma^2}\left( \frac{1}{\sigma^2}\right) = 0 $$ 
$$ \implies -\frac{n}{2\sigma^2}-\left[\frac{1}{2}\sum_{i=1}^{n}(X_i-\mu)^2 \right]\left(-\frac{1}{(\sigma^2)^2} \right)= 0 $$
$$ \implies -\frac{n}{2\sigma^2}+\left[\frac{1}{2}\sum_{i=1}^{n}(X_i-\mu)^2 \right]\frac{1}{(\sigma^2)^2} = 0 $$
$$ \frac{1}{2\sigma^2}\left[\frac{1}{\sigma^2}\sum_{i=1}^{n}(X_i-\mu)^2-n \right] = 0 $$

which, if we rule out $\sigma^2=0$, is equal to zero only if, for some $\hat{\sigma}^2$, 

$$ \boxed{\hat{\sigma}^2 = \frac{1}{n}\sum_{i=1}^{n}(X_i-\hat{\mu})^2} $$
\newpage


\section{2.}

Let $X_1, X_2, \dots ,Xn$ be i.i.d. samples from Poisson ($\lambda$). Then Likelihood is defined as
$$ L(X_1, X_2,\dots,X_n|\lambda) = \prod_{i=1}^{n} e^{-\lambda}\frac{1}{X_i!}\lambda^{X_j} $$

For simplicity, we consider logarithm. Taking log with base $e$ on both sides and simplifying, we get

$$ l(X_1,X_2,\dots,X_n|\lambda) := ln(L(X_1, X_2, \dots, X_n| \lambda)) $$
$$ = -n\lambda -\sum_{i=1}^{n}ln(X_i!) +ln(\lambda)\sum_{i=1}^{n}X_i $$

Maximum likelihood estimate for $\lambda$ is found by differentiating the expression w.r.t. to the mean ($\lambda$).

Let
$$ \frac{d}{d\lambda}l(X_1,X_2,\dots,X_n|\lambda) = 0 $$
$$ \implies \frac{d}{d\lambda} \left(-n\lambda -\sum_{i=1}^{n}ln(X_i!) +ln(\lambda)\sum_{i=1}^{n}X_i \right) = 0 $$
$$ \implies -n +\frac{1}{\lambda}\sum_{i=1}^{n}X_i = 0 $$
$$ \implies \boxed{\hat{\lambda} = \frac{1}{n}\sum_{i=1}^{n}X_i} \quad \text{for some} \; \hat{\lambda}$$ 
\newpage


\section{3.}

\subsection{(0)}
The method used to calculate 'Nhat' and 'phat' is method of moments.
This is done by equating
$$ \mu_k(X_1,X_2,\dots,X_n) = \mathbb{E}(X^k) = m_k(X_1,X_2,\dots,X_n) $$
where 
$$ \mu_{k}(X_1,X_2,\dots,X_n) = \frac{\sum_{i=1}^{n}{X_i}^k}{n} $$

Since the distribution is binomial,
$$ \mu_1 = \mathbb{E}(X) = Np \qquad \text{and} $$
$$ \mu_2 = \mathbb{E}(X^2) = Var(X) + [\mathbb{E}(X)]^2 = Np(1-p) + (Np)^2 $$

Hence, $m_1 = Np$ and $m_2 = Np(1-p)+(Np)^2$
$$ \therefore \; m_2 = m_1(1-p) + m_1^2 $$
$$ \implies \hat{p} = \frac{m_1-(m_2-m_1^2)}{m_1} $$
and then, 
$$ \hat{N}=\frac{m_1}{\hat{p}}=\frac{m_1^2}{m_1-(m_2-m_1^2)} $$
We get the estimated values of the quantities N, p as $\hat{N}$ and $\hat{p}$.
\newpage

\subsection{(a)}
```{r}
n <- 100
N <- 20
p <- 0.4
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```

\subsection{(b)}
```{r}
n <- 1000
N <- 20
p <- 0.4
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```

The error here, with more data, is significantly lower.
\newpage

\subsection{(c)}
```{r}
n <- 100
N <- 10
p <- 0.1
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```

```{r}
n <- 100
N <- 10
p <- 0.9
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```
\newpage

```{r}
n <- 1000
N <- 10
p <- 0.1
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```

```{r}
n <- 1000
N <- 10
p <- 0.9
x <- rbinom(n, N, p)
m1 <- mean(x)
m2 <- mean(x^2)
Nhat <- m1^2 / (m1 -(m2-m1^2))
phat <- (m1 - (m2-m1^2)) / m1
```

```{r}
Nhat -N
phat -p
```
\newpage


\section{4.}
Let $M$ be a random variable, defined as \
$$ M:= \text{maximum number of heads obtained} $$
and $P[M=r]$ and $F_M(r)$ denote, respectively, the PMF and CDF for this random variable. 

\subsection{(a)}
We need the value of $P[M=30]$. \
Since all the $n$ ($n$=$10$) trials are independent and identical (as i.i.d.s), 
$$ F_M(r) = P[M \leq r] = P[X_i \leq r]^{10} $$
and $$ P[X_i \leq r] = \sum_{k=0}^r \binom{50}{k}p^k(1-p)^{50-k} $$
$$ \therefore F_M(r) = \left[\sum_{k=0}^r \binom{50}{k}p^k(1-p)^{50-k}\right]^{10} $$
Now,
$$ P[M=30] = F_M(30) - F_M(29) $$
$$ = \left[\sum_{k=0}^{30} \binom{50}{k}p^k(1-p)^{50-k}\right]^{10} - \left[\sum_{k=0}^{29} \binom{50}{k}p^k(1-p)^{50-k}\right]^{10} $$

\subsection{(b)}
When we want to find an estimate for $p$ we usually take the probability function and differentiate it w.r.t. $p$, equate it with $0$. In this case tho, it is a $500$ degree polynomial and differentiating it gives us a $499$ degree polynomial which is unsolvable by us under 5 mins.

Nevertheless, we can try to provide a (good enough) estimate.
Defining $h_i$ as the number of heads in the $i^{th}$ trial. \
$\therefore$ the empirical probability of getting heads in $i^{th}$ trial
$$ P_{h_i}= \frac{h_i}{50} $$

So, estimating $p$ (probability of getting head in each toss) as the mean of these probabilities is my best guess.
$$ p = \frac{\sum_{i=1}^{10}P_{h_i}}{10}= \frac{\sum_{i=1}^{10}h_i}{500} $$
\newpage

\subsection{(c)}
```{r}
likelihood_max <- function(p, M=30, N=50, n=10) {
pbinom(q = M, size = N, prob = p)^n - pbinom(q = M - 1, size = N, prob = p)^n
}
```
According to the graph, the best estimate for 'p' is at about neighbourhood of 0.5, where it attains the maximum value.

\subsection{(d)}
\subsubsection{i.}
```{r fig.height=4, fig.width=6}
plot(likelihood_max, from=0, to =1)
```

\subsubsection{ii.}
```{r}
optimize(likelihood_max, interval=c(0,1), maximum=TRUE)
```
