---
title: '**Introduction to Statistics and Computation with Data: AppleBread Problem Solution 2**'
author: "Daibik Barik (Bmat2316)"
date: "2024-03-22"
output:
  html_document:
    citation_package: pandoc-citeproc
    number_sections: TRUE
    df_print: paged
header-includes: \usepackage{bm, amsmath, amssymb}
bibliography: ["Reference.bib"]
link-citations: TRUE
---

```{r}
#set.seed(2024)
library(tidyverse)
library(viridis)
library(ggplot2)
library(gganimate)
library(gifski)
```
\newpage


<font size="8"> **Challenge Problem 2]** </font> 

```{r}
##Functioning of the model
process<- function(A, K) {
  #Creation
  for(i in 1:K) {
    boundary= c()
    for(element in A) {
      if(! (element +1) %in% A) {
        boundary<- append(boundary, (element+1))
      }
      if(! (element -1) %in% A) {
        boundary<- append(boundary, (element-1))
      }
    }
    A<- append(A, sample(boundary, 1))
  }
  
  #Annihilation
  for(i in 1:(K-1)) {
    length= length(A)
    remove_index= sample(1:length, 1)
    A<- A[-c(remove_index)]
  }
  return(A)
}
```

# .
Let the number of elements of $A_n$ be $N$ at time $n$.

At time $n+1$, $K$ new sites are added one by one, making the total number of occupied sites $N+K$. After that $K-1$ sites are chosen uniformly from those $N+K$ sites and are deleted, one by one.

Therefore, in the end, the total number of sites result to $N+K-(K-1)= N+1$.
\newpage

# .
```{r}
##Code Run for 100 steps
A= c(0); K= 3
A_n= list(A)
for(n in 1:100) {
  A= process(A, K)
  A_n[[n+1]]= A 
}
#print(length(A)) for verification
```

As per the discussion with Siva in class of $22^{nd}$ March $2024$, this model is being represented for $n=100$ instead of $n=1000$, as in the second case the plot is indistinguishable from a straight line. 

Siva also suggested the use of "Gaps" instead of the "Occupied Sites" to one of our fellow students. But personally I didn't like the plot at all since it didn't help to provide an intuition, or what was in this case, not a proper simlation of the Model.  

```{r}
##Static Representation of Model
x_values= c(0)
for(i in 1:100) {
  x_values= c(x_values, rep(i,i+1))
}
A_n_flattened= unlist(A_n, use.names= FALSE)
df= data.frame(
  x_cords= x_values,
  y_cords= A_n_flattened 
)
ggplot(data= df) +
  geom_point(mapping= aes(x= x_cords, y= y_cords),
             size= 1/2) +
  geom_line(mapping= aes(x= x_cords, y= 0), col= "darkred")+
  xlab("n Values") + ylab("A_n sites") +
  labs(title= "The Annhilation Creation model", 
       subtitle= "Static Sites Representation" )
```

```{r}
##Dynamic Simulation of model
model= ggplot(data= df) +
  geom_point(mapping= aes(x= x_cords, y= y_cords),
             size= 1/2) +
  transition_manual(x_values)+ 
  xlab("n Values") + ylab("A_n sites") +
  labs(title= "The Annhilation Creation model", 
       subtitle= "Dynamic Simulation" )
animate(model, fps= 7.5)
```
\newpage

# .
```{r}
##Functioning of L(A_n) and G2(A_n)
L<- function(A) {
  A_sorted= sort(A)
  A_gaplength= diff(A_sorted)-1
  sum(A_gaplength)
}

G2<- function(A) {
  A_sorted= sort(A)
  A_gaplength= diff(A_sorted)-1
  sum(A_gaplength >=2)
}
```

```{r}
##Finding L_n and G2_n for n=1000 over 100 trials
L_tn= matrix(nrow= 100, ncol= 1000)
G2_tn= matrix(nrow= 100, ncol= 1000)
trials= 1:100; n_values= 1:1000
for(t in trials) {
  A= c(0); K= 3
  for(n in n_values) {
    A= process(A, K)
    L_tn[t,n]= L(A)
    G2_tn[t,n]= G2(A)
  }
}
```
\newpage
From theorem $2.1.1.$ [@Thesis], we say $\lim sup_{n\to\infty} \frac{L(A_n)}{h(n)}= \frac{K-1}{K}$, which means $\exists\; N_0 \in \mathbb{N}$ s.t. $L(A_n)$ and thus $\overline{L(A_n)}= L_n\leq \frac{2}{3}h(n)\; \forall\; n\geq N_0$ as, in this case $K=3$.

We try to plot the line graph of $\frac{L(A_n)}{h(n)}$, and also show the line $y=\frac{2}{3}$, under which the graph should lies $\forall\; n\geq N_0$.

```{r}
##Showing (by plot) that L_n= O(h(n)) 
L_n= c()
for(n in n_values) {
  L_n[n]= mean(L_tn[,n])
}
h<- function(n) {
  log(n)/ log(log(n))
}
n_val<- 1:1001
plot(n_val, c(0, L_n/h(n)),
     xlab= "n Values", ylab= "L_n/ h(n)",
     main= "The Upper Bound for L_n/h(n)", type='l')
y_lim<- 0*n_val +2/3
lines(n_val, y_lim, col = "darkred", lty = 2)
text(x= 5, y= 7/5,
     labels= "y= 3/2", col= "darkred")
```

\newpage

We try to show, as Theorem $2.1.2$ [@Thesis], that 
$$ \mathbb{P}\left[G^{(2)}(A_n)= 1\; \text{ i.o.}\right]= 1 $$
$$ \implies \mathbb{P}\left[\;\forall N\geq 1\; \exists n\geq N\; s.t.\; G^{(2)}(A_n)= 1\right]= 1 $$
$$ \implies \mathbb{P}\left[\;\forall N\geq 1\; \exists n\geq N\; s.t.\; \overline{G^{(2)}(A_n)}= G^{(2)}_n(A_n)= 1\right]= 1 $$

Trying to find the empirical probability through statistics....


```{r}
##Showing by plot P[G2_n(A_n)=1 i.o.]= 1

pG2_1= matrix(nrow= 100, ncol= 1000)
for(t in trials) {
  max_g2n_1= 1
  for(n in n_values) {
    if(G2_tn[t, n]== 1) {
      max_g2n_1= n
    }
  }
  for(n in n_values) {
    if(n<= max_g2n_1) {
      pG2_1[t,n]= 1
    } else {
      pG2_1[t,n]= 0
    }
  }
}
pG2n_1= c()
for(n in n_values) {
  pG2n_1[n]= mean(pG2_1[,n])
}
#print(pG2n_1)
plot(n_val, c(1, pG2n_1),
     xlab= "n Values", ylab= "P[N: N>=n, G2(N)=1]",
     main= "Probability of Existence of Gap-2", type = "l")
```

We see for lower values of $n$, $\mathbb{P}= 1$. But as we go higher, the probability decreases since there are limited number of samples taken into consideration. 

Therefore, if we take the whole Natural Number line, we can say $\mathbb{P}\left[G^{(2)}(A_n)= 1\; \text{ i.o.}\right]= 1$

`r if (knitr::is_html_output()) '# References {-}'`














