---
title: "**Introduction to Statistics and Computation with Data: HomeWork 5**"
author: "Daibik Barik (Bmat2316)"
date: "2024-02-28"
output: pdf_document
---

```{r}
set.seed(2024)
```

\section{1.}
```{r}
E= rexp(100)
qqnorm(E);qqline(E)
```

It follows the Normal distribution from the interval $(-1,1)$ and positively deviates from the normal graph on both the tails.
\newpage

```{r}
U= runif(100)
qqnorm(U);qqline(U)
```

It follows the Normal distribution from the interval $(-1,1)$ and positively deviates from the normal graph on the front tail and negatively deviates in the end tail.
\newpage


\section{2.}

```{r}
SupDF = read.csv(file= "./SuppandiZoomLog.csv", header = TRUE)
library(tidyverse)
#SupDF
```
\newpage

\subsection{(a)}
```{r}
ggplot(data= SupDF) +
  geom_bar(mapping= aes(x=Venue, fill = Meeting.Topic)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
\newpage

\subsection{(b)}
```{r}
ggplot(data= SupDF, mapping= aes(x= Meeting.Topic, y= Duration..Minutes.)) +
  geom_boxplot() + geom_jitter(mapping= aes(colour= Meeting.Topic))
```
\newpage

\subsection{(c)}
```{r}
library(lubridate)
SupDF$Day= weekdays(as.Date(SupDF$Start.Time))
#SupDF
SupWeek= subset(SupDF, Day!= "Sunday" & Day!= "Saturday")
ggplot(data= SupWeek) + 
  geom_bar(mapping= aes(x= Day), fill= "lightblue", colour= "black") +
  coord_polar() 
```
\newpage

\subsection{(d)}
```{r}
SupDF$Start.Hour= hour(SupDF$Start.Time)
ggplot(data= SupDF) +
  geom_histogram(mapping= aes(x= Start.Hour), fill="lightblue", color="black",
                 binwidth=1)
```
\newpage

\subsection{(e)}
\paragraph{Null Hypothesis$(H_0)$:} The meeting time(Hours) follows exponential distribution with parameter = $\frac{1}{14}$
\paragraph{Alternative Hypothesis$(H_1)$:} The meeting time(Hours) does not follow exponential distribution with parameter = $\frac{1}{14}$
```{r}
mean(SupDF$Start.Hour)
ks.test(SupDF$Start.Hour, 'pexp', rate = 1/14)
```
As the \emph{p-value} is less than the level of significance, so we accept the $H_1$ and reject the $H_0$.
\newpage

\section{3.}
```{r}
x = c(1,2,3,4,5,6)
probx= c(1/6,1/6,1/6,1/6,1/6,1/6)
Rolls=sample(x, size=1500, replace=T, prob=probx)
Rollm=matrix(Rolls, nrow = 5)
 Rollsums = apply(Rollm, 2, sum)
library(ggplot2)
density = function(x,a,s){ (1/((2*pi)^(0.5)*s ))* exp(-(x-a)^2/(2*s^2))}
dfrolls = data.frame(Rollsums)
mu = mean(dfrolls$Rollsums)
sigma= sd(dfrolls$Rollsums)
ggplot(data=dfrolls) + 
  geom_histogram(mapping=aes(x=Rollsums,y=after_stat(density)),
                 color="#00846b", fill=NA, binwidth=1) +
  xlim(5,30) + geom_function(fun=density,args=list(a=mu, s= sigma),
                             color="black")
```
\newpage

\subsection{(a)}
* The **matrix()** function is utilized to organize a set of values into a matrix format. Here, the *matrix()* function is used to transform the 1500 generated samples(Rolls) into a matrix with 5 rows and 300 columns. 

* The **apply()** function computes a vector, array, or list of values by applying a specified function to the margins (rows or columns) of an array or matrix. Here, *apply()* calculates the sum of elements along row 2 (margin) and stores 300 sums.

\subsection{(b)}
The most approximate value of $\int_{12}^{21}\text{density}(x,mu,sigma)dx$ is the area under the histogram from $x=12$ to $x=21$.
```{r}
area_hist= 0
for(x in 12:21)
  area_hist= area_hist + sum(Rollsums== x)
area_hist/300
```

Hence the approximate area is $0.773$.

\subsection{(c)}
As the histogram follows fairly normal distribution, in the area under the graph, to approximate $a$ and $b$ by standardizing $12$ and $21$, we can use the formula for standardization:
$$ z=\frac{x-\mu}{\sigma} $$
where $\mu$ and $\sigma$ are-
```{r}
mu;sigma
```
$$ \therefore \; a=\frac{12-17.83667}{3.908227} \quad \text{and} \quad b=\frac{21-17.83667}{3.908227} $$
```{r}
a= (12-mu)/sigma
b= (21-mu)/sigma
a;b
```
$$ \implies a= -1.493431 \quad \text{and} \quad b= 0.8094037 $$
which are more or less approximate guesses.
\newpage

\subsection{(d)}
```{r}
Dice= read.csv(file="Dice.csv",head=TRUE)
names(Dice)= "Sum";
Dice$Sum= as.numeric(Dice$Sum);
Dice= na.omit(Dice$Sum);mu=mean(Dice);sigma=sd(Dice);
density= function(x,a,s){ (1/((2*pi)^(0.5)*s ))* exp(-(x-a)^2/(2*s^2))}
diceroll= data.frame(Dice)
ggplot(data= diceroll) +
  geom_histogram(mapping= aes(x= Dice, y= after_stat(density)),
                 color= "#00846b", fill= NA, binwidth= 1) +
  xlim(5,30) + geom_function(fun= density, 
                             args= list(a= mu, s= sigma),
                             color="black") +
  xlab("Sum of Rolls")
```
\newpage

$\int_{12}^{21}\text{density}(x,mu,sigma)dx$ can be approximated by calculating the area of the histogram between $x=12$ and $x=21$.

```{r}
area_hist= 0
for(x in 12:21)
  area_hist= area_hist + sum(Dice== x)
area_hist / length(Dice)
```
Hence the approximate area is $0.807$.

To approximate $a$ and $b$ by standardizing $12$ and $21$, we can use the formula for standardization:
$$ z=\frac{x-\mu}{\sigma} $$
where $\mu$ and $\sigma$ are-
```{r}
mu;sigma
```
$$ \therefore \; a=\frac{12-17.23077}{3.670899} \quad \text{and} \quad b=\frac{21-17.23077}{3.670899} $$
```{r}
a= (12-mu)/sigma
b= (21-mu)/sigma
a;b
```
$$ \implies a= -1.424929 \quad \text{and} \quad b= 1.026787 $$
which are more or less approximate guesses.
\newpage


\section{4.}

\subsection{(a)}
```{r}
x=seq(-2,2,0.1)
y=pnorm(x)
plot(x,y,type='l')
```
\newpage

\subsection{(b)}
```{r}
x=seq(-2,2,0.1)
m=100
p=0.4
y=pbinom(x*((m*p)*(1-p))^(1/2)+m*p,m,p)
plot(x,y, type="l")
```
\newpage

\subsection{(c)}
```{r}
m= 100; p= 0.4;
Bin= rbinom(1000,m,p)
density= function(x,a,s){(1/((2*pi)^(0.5)*s ))*exp(-(x-a)^2/(2*s^2))}
mu= m*p; sigma= (m*p*(1-p))^(1/2);
BinDf= data.frame(Bin)
ggplot(BinDf) + geom_histogram(mapping=aes(x=Bin,y=after_stat(density)),
color="#00846b", fill=NA, binwidth=1)+
geom_function(fun=density,args=list(a=mu, s= sigma), color="black")+
xlab("Sample Outcome")
```

\paragraph{Conclusion:} As $m$ increases, the binomial distribution converges to Normal distribution with the parameters- mean($\mu$): $mp$, variance($\sigma^2$): $mp(1-p)$.










