---
title: "**Introduction to Statistics and Computation with Data: HomeWork 3**"
author: "Daibik Barik (Bmat2316)"
date: "2024-01-31"
output: pdf_document
---

\section{1}
```r
install.packages("forcats")                                                
install.packages("gapminder")
```

```{r setup, echo=FALSE}
install.packages(c("gapminder", "forcats"),
  repos = "http://cloud.r-project.org"
)
```
\newpage

\section{2}
```{r}
library(tidyverse)
library(gapminder)
str(gapminder$continent)
```

Concisely exhibit the internal composition of an R object using a diagnostic function, serving as an alternative to the *summary()*.
So here it displays about the 5 factors of *gapminder*, which are also continents named "Africa", "America" etc.

```{r}
levels(gapminder$continent)
```

The *levels()* grants access to the levels attribute of a variable. Here the levels are named after different continents like "Africa", "Americas", etc.

```{r}
nlevels(gapminder$continent)
```

Return the number of levels which its argument has. Here it has exactly 5.

```{r}
class(gapminder$continent)
```

Several R objects contain a class attribute, which is a character vector that provides the names of the classes from which the object inherits.
\newpage

\section{3}
```{r}
gapminder %>% count(continent)
```

*count()* lets you quickly count the unique values of frequencies of one or more variables.

```{r}
fct_count(gapminder$continent)
```

*fct_count()* Count entries in a factor, here it gives same output because number of entries in factors and frequencies of the variables are same.
\newpage

\section{4}
```{r}
nlevels(gapminder$country)
```

Prints the number of levels in the country column of the gapminder dataset.

```{r}
h_countries <- c("Egypt", "Haiti", "Romania", "Thailand", "Venezuela")
h_gap <- gapminder %>% filter(country %in% h_countries)
nlevels(h_gap$country)
```

It prints the number of levels in the country column, and **since the other levels are still present with 0 entries**, it shows the total number of levels same as before.

```{r}
h_gap_dropped <- h_gap %>%droplevels()
nlevels(h_gap_dropped$country)
```

*droplevels()* removes the unused levels from a factor. So, now if we count number of levels, it returns 5.

```{r}
h_gap$country %>% fct_drop() %>%levels()
```

*fct_drop()* drops the unused levels in the factor, and *levels()* print the remaining levels, by default in alphabetical order.
\newpage

\section{5}
```{r}
gapminder$continent %>% levels()
```
Prints the levels in the continent column of Gapminder dataset.

```{r}
gapminder$continent %>% fct_infreq() %>% levels()
```

Prints the levels in the decreasing order of the number of observations.

```{r}
gapminder$continent %>% fct_infreq() %>%fct_rev() %>%levels()
```

Prints the levels in the increasing order of the number of observations (**opposite to the order made by *fct_infreq***)
