---
title: "**Introduction to Statistics and Computation with Data: HomeWork 9**"
author: "Daibik Barik (Bmat2316)"
header-includes:
    - \usepackage{bm, amsmath, amssymb}
date: "2024-03-23"
output: pdf_document
---

\section{1.}

\subsection{(a)}
Let $x$ be the mean household income.
\paragraph{Null Hypothesis: } $x= 50000$
\paragraph{Alternate Hypothesis: } $x\neq 50000$

\subsection{(b)}
Let $t$ be Siva's mean commute time to ISI in minutes.
\paragraph{Null Hypothesis: } $t\geq 90$
\paragraph{Alternate Hypothesis: } $t <90$

\subsection{(c)}
Let $t$ be the mean delay arrival time in minutes.
\paragraph{Null Hypothesis: } $t\leq 15$
\paragraph{Alternate Hypothesis: } $t >15$

\subsection{(d)}
Let $p$ be the chance(or probability) that superstudent will submit homework late.
\paragraph{Null Hypothesis: } $p <0.15$
\paragraph{Alternate Hypothesis: } $p\geq 0.15$
\newpage


\section{2.}
Given $X_1, X_2, \dots, X_{100} \sim \mathcal{N}(\mu, 3)$.
We use the $Z$-test for testing of the hypothesis. 

(Significance level of $\alpha= 0.05$)

\paragraph{Null Hypothesis: } $\mu= c$
\paragraph{Alternate Hypothesis: } $\mu> c$

\subsection{(a)}
Consider $Y_1, Y_2, \dots, Y_{100} \sim^{iid} \mathcal{N}(8.5, 3)$ under the null hypothesis ($\mu= 8.5$). 

By CLT, we know, $$ \frac{\sqrt{n}(\overline{Y}- \mu_Y)}{\sigma_Y} \sim \mathcal{N}(0,1)= Z\; (\text{let}) $$
$\begin{aligned}
\mathbb{P}[\overline{Y} \geq \overline{X}] &= \mathbb{P}[\overline{Y} \geq 10.2] \\
&= \mathbb{P}\left[\frac{\sqrt{n}(\overline{Y}- \mu_Y=\mu)}{\sigma_Y} \geq \frac{\sqrt{n}(10.2- \mu_X=\mu)}{\sigma_X}\right] \\
&= \mathbb{P}\left[\frac{\sqrt{100}(\overline{Y}- 8.5)}{\sqrt{3}} \geq \frac{\sqrt{n}(10.2- 8.5)}{\sqrt{3}}\right] \\
&= \mathbb{P}\left[Z \geq \frac{10\times (1.7)}{1.732}\right] \\
&= \mathbb{P}\left[Z \geq \frac{17}{1.732}= 9.81\right]
\end{aligned}$
```{r}
1-pnorm(9.81)
```
Therefore, $\mathbb{P}\left[\overline Y \geq 10.2\right]= 0$, which is less than our $\alpha= 0.05$ and hence we reject our null hypothesis and accept our.

\subsection{(b)}
Consider $Y_1, Y_2, \dots, Y_{100} \sim^{iid} \mathcal{N}(9.5, 3)$ under the null hypothesis ($\mu= 8.5$).

By CLT, we know, $$ \frac{\sqrt{n}(\overline{Y}- \mu_Y)}{\sigma_Y} \sim \mathcal{N}(0,1)= Z\; (\text{let}) $$
$\begin{aligned}
\mathbb{P}[\overline{Y} \geq \overline{X}] &= \mathbb{P}[\overline{Y} \geq 10.2] \\
&= \mathbb{P}\left[Z \geq \frac{7}{1.732}= 4.04\right]
\end{aligned}$
```{r}
1-pnorm(4.04)
```
Therefore, $\mathbb{P}\left[\overline Y \geq 10.2\right]\approx 2.67256\times 10^{-5}$, which is less than our $\alpha= 0.05$ and hence we reject our null hypothesis.
\newpage


\section{3.}

Te random variables in question 







