---
title: "**Introduction to Statistics and Computation with Data: HomeWork 1**"
author: "**Daibik Barik**"
date: "2024-01-15"
output: pdf_document
---

\section*{Q2.}
\subsection*{(a)}
Using the command 'diff':-
```{r}
instasee= c(8,8,3,12,13,10,1,8,7)
timediff= diff(instasee)
timediff
```
As we can see, 'diff' gives us the difference between the consecutive values in the vector. 

\subsection*{(b)}
Finding the maximum, minimum and average time between two logins:-
```{r}
x = timediff %% 24 #Taking remainder with 24 to get the number of hours between two logins
x
x[1] = 24 
max(x)
mean(x)
min(x)
```
\section*{Q3.}
\subsection*{(a)}
Finding the maximum, minimum and average scores:-
```{r}
scoreSS = c(7, 6, 10, 8, 7, 9, 9, 6, 4, 10, 8, 6, 9, 10)
max(scoreSS)
mean(scoreSS)
min(scoreSS)
```
\subsection*{(b)}
Correcting the incorrect entry 4 to '5':- 
```{r}
scoreSS_edit = scoreSS
scoreSS_edit[4] = 5
mean(scoreSS_edit)
```
\subsection*{(c)}
Using the command sum with a condition:-
```{r}
sum(scoreSS >= 9)
```
\subsection*{(d)}
The above command in part (c) gives the number of scores that are greater than or equal to 9.  

Noting by observation, all the scores are less than 17. Following is the procedure to get the percentage of scores less than 17 using R:-
```{r}
n = sum(scoreSS < 17)
n
percent = n/length(scoreSS) * 100
percent
```
\section*{Q4.}
\subsection*{(a)}
Finding the total expenditure on phone bills in the whole year:-
```{r}
Shreelakshmibill = c(460, 330, 390, 370, 460, 300, 480, 320, 490, 350, 300, 480)
sum(Shreelakshmibill)
```
\subsection*{(b)}
Finding the minimum and maximum bill amounts:-
```{r}
min(Shreelakshmibill)
max(Shreelakshmibill)
```
\subsection*{(c)}
Percentage of bill amounts greater than 400:- 
```{r}
n = sum(Shreelakshmibill > 400)
n
percent = n/length(Shreelakshmibill) * 100
percent
```
\subsection*{(d)}
Calculating the available balance per month and its average value if the total money is 3000 every month:-
```{r}
freemoney = 3000 - Shreelakshmibill
freemoney
mean(freemoney)
```
\section*{Q5.}
\subsection*{(a)}
The dataset diamonds contains details of round cut diamonds:-  
a) A dataset containing the prices and other attributes of almost 54,000 diamonds.
b) Has different variables like price, carat, cut, color, depth, clarity, x, y, z, table.


\subsection*{(b)}
```{r, include = FALSE}
library(tidyverse)
library(viridis)
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping=aes(x = cut, fill = clarity),
                                   position = "dodge") +  scale_color_viridis_d()
```
\subsection*{(c)}
Observation from graph :-  
1. As the cut progresses from far to ideal, the average count of all the clarities increases. This can be observed from the height of the histograms along the Y-axis.